---
title: "Project setup"
site: workflowr::wflow_site
output:
  workflowr::wflow_html:
    toc: true
---

```{r knitr-opts, include=FALSE}
knitr::opts_chunk$set(comment = "#",collapse = TRUE)
```

**Last updated:** `r Sys.Date()`

**Code version:** `r system("git log -1 --format='%H'", intern = TRUE)`

**To reproduce the results from this project, please follow these instructions.**

*In general, `drake` was used to manage long-running code and `workflowr` was used to manage the [website](https://jennysjaarda.github.io/PSYMETAB/index.html).*

***

## Step 1: initiate project on remote server.

All processing scripts were run from the root sgg directory.
Project was initialized using `workflowr` rpackage, see [here](https://jdblischak.GitHub.io/workflowr/articles/wflow-01-getting-started.html).

On sgg server:

```{r eval=FALSE}
project_name <- "PSYMETAB"
library("workflowr")

wflow_start(project_name) # creates directory called project_name

options("workflowr.view" = FALSE) # if using cluster
wflow_build() # create directories
options(workflowr.sysgit = "")

wflow_publish(c("analysis/index.Rmd", "analysis/about.Rmd", "analysis/license.Rmd"),
              "Publish the initial files for myproject")

wflow_use_GitHub("jennysjaarda") # select option 2: manually create new repository

wflow_git_push()
```

You have now successfully created a GitHub repository for your project that is accessible on GitHub and the servers.

Next setup a local copy.

## Step 2: Create local copy on personal computer.

Within terminal of personal computer, clone the git repository.

 ```bash
cd ~/Dropbox/UNIL/projects/
git clone https://GitHub.com/jennysjaarda/PSYMETAB.git PSYMETAB
 ```

Open project in atom (or preferred text editor) and modify the following files:

- Because Terminal cannot generate a preview and `workflowr` doesn't like the sysgit, to the `.Rprofile` file, add:
  - `options(workflowr.sysgit = "")`
  - `options("workflowr.view" = FALSE)`
- To ensure git hub isn't manaaging large files, modify the `.gitignore` file, by adding the following lines:
  - `analysis/*`
  - `data/*`
  - `!analysis/*.Rmd`
  - `!data/*.md`
  - `.git/`
- Save and push these changes to GitHub.
- Pull to the server.

## Step 3: Create internal project folders.

Return to sgg server and run the following:

```bash
project_dir=/data/sgg2/jenny/projects/PSYMETAB
mkdir $project_dir/data/raw
mkdir $project_dir/data/processed
mkdir $project_dir/data/raw/reference_files
mkdir $project_dir/data/raw/phenotype_data
mkdir $project_dir/data/raw/extraction
mkdir $project_dir/data/processed/phenotype_data
mkdir $project_dir/data/processed/extraction
mkdir $project_dir/docs/assets
```

This will create the following directory structure in `PSYMETAB/`:

```
PSYMETAB/
├── .gitignore
├── .Rprofile
├── _workflowr.yml
├── analysis/
│   ├── about.Rmd
│   ├── index.Rmd
│   ├── license.Rmd
│   └── _site.yml
├── code/
│   ├── README.md
├── data/
│   ├── README.md
│   ├── raw/
|       ├── phenotype_data/
|       ├── reference_files/
|       └── extraction/
│   └── processed/
|       ├── phenotype_data/
|       ├── reference_files/
|       └── extraction/
├── docs/
|       └── assets/
├── myproject.Rproj
├── output/
│   └── README.md
└── README.md
```

Raw PLINK (`ped` and `map` files) data were copied from the CHUV folder (`L:\PCN\UBPC\ANALYSES_RECHERCHE\Jenny\PSYMETAB_GWAS`) after being built in genomestudio to the `data/` drive.

## Step 4: Run and summarize analyses.

### Part A: Setup drake and execute plan.

**Note that Part A and B are happening in parallel.**

For execution of drake plan: see [`make.R`](https://GitHub.com/jennysjaarda/PSYMETAB/blob/master/make.R)
For drake plan see: [`code/plan.R`](https://GitHub.com/jennysjaarda/PSYMETAB/blob/master/code/plan.R)

Configure a slurm template
```{r eval=FALSE}
options(clustermq.scheduler = "slurm", clustermq.template = "slurm_clustermq.tmpl")
drake_hpc_template_file("slurm_clustermq.tmpl")

# Write the file slurm_clustermq.tmpl. and edit manually
```
The file created using the `clustermq` template was edited manually to match [`slurm_clustermq.tmpl`](https://GitHub.com/jennysjaarda/PSYMETAB/blob/master/slurm_clustermq.tmpl)

```{r comment=''}
cat(readLines('slurm_clustermq.tmpl'), sep = '\n')
```

### Part B: Build and maintain website.

Follow the general workflow outlined by [workflowr](https://jdblischak.GitHub.io/workflowr/articles/wflow-01-getting-started.html#the-workflow), with some minor revisions to accomodate workflow between personal computer and remote server:

  1. Open a new or existing R Markdown file in `analysis/` (optionally using `wflow_open()`). (Usually created manually on personal computer and push to server to build later.) If creating manually, add the following to the top of the R Markdown file with an appropriate name for `Title`:

```{r eval=FALSE}
---
title: "Title"
site: workflowr::wflow_site
output:
  workflowr::wflow_html:
    toc: true
---
```

  2. Write documentation and perform analyses in the R Markdown file.

  3. Run `commit` and `push` to upload revised R Markdown file to GitHub repository.

  4. On server, pull changes using `wflow_git_pull()` (optionally using `git pull` from Terminal within cloned repository).

  5. Within R console, run `wflow_build()`. This will create html files with `docs/` folder. These files cannot be viewed directly on server, but can be transfered and viewed via [FileZilla](https://filezilla-project.org/) or viewed directly by mounting the remote directory to your personal computer using [SSHFS](https://en.wikipedia.org/wiki/SSHFS) (recommended).

  6. Return to step 2 until satisfied with the result (optionally, edit Rmd file directly on server using `vi` if only small modifications are necessary).

  7. Run `wflow_status()` to track repository.

  8. Run `wflow_publish()` to commit the source files (R Markdown files or other files in `code/`, `data/`, and `output/`), build the HTML files, and commit the HTML files. If there are uncommited files in the directory that are not ".Rmd", `wflow_publish(all=T)` does not work. Alternatively, run the following with an informative `message`:

```{r eval=FALSE}
repo_status <- wflow_status()
rmd_commit <- c(rownames(repo_status$status)[repo_status$status$modified],
         rownames(repo_status$status)[repo_status$status$unpublished],
         rownames(repo_status$status)[repo_status$status$scratch])

wflow_publish(rmd_commit,
              message="Updating webiste")

```

  9. Push the changes to GitHub with` wflow_git_push()` (or `git push` in the Terminal).
